### 2015-03-20 cug_paper/01_introduction.tex
Add directory to hold the CUG paper

\section{Introduction}
\label{sec:intro}

Discuss the general idea and the rationale behind it