### 2015-03-20 cug_paper/02_design.tex

\section{Design And Implementation}
\label{sec:design}

Discuss how the software works.  Things like using CCI for small messages, bulk transfers happening by way of cudaMemcpy().  Mention Cray-specific stuff such as setting 'feature=gpudefault' in the qsub script.