### 2015-03-20 cug_paper/03_analysis.tex

\section{Analysis}
\label{sec:analysis}

\subsection{Experiment Setup}
Discuss what we ran, the parameters we swept and why we chose those parameters

\subsection{Results}
Discuss the results