### 2015-03-20 cug_paper/04_futurework.tex

\section{Future Work}
\label{sec:future}

Discuss work yet to be done:  add support for caching in system memory,  make 'production ready' (ie: easier for applications to use)