### 2015-03-20 cug_paper/05_conclusion.tex

\section{Conclusion}
\label{sec:conclusion}

Discuss what we learned; summarize