### 2015-03-27 cug_paper/04_futurework.tex

\section{Future Work}
\label{sec:future}

\todo[inline]{Discuss work yet to be done:  add support for caching in system memory,  make 'production ready' (ie: easier for applications to use), measure/quantify any negative impacts on app performance (particularly for apps that run 16 ranks/node)}

\todo[inline]{ Summit will have node-local SSD's and that a variant of this code could be quite useful for applications that haven't been modified to take advantage of them more directly.  Not sure if we can mention Summit details yet.  Not sure if we should mention Summit at a Cray conference...  Maybe this paragraph should be in terms of generic burst-buffer support.}