### 2015-03-27 cug_paper/05_conclusion.tex

\section{Conclusion}
\label{sec:conclusion}

\todo[inline]{Discuss what we learned; summarize}